Coupling PDEs on 3D-1D domains with Lagrange multipliers

### 2020-03-14 3d1d_coupled_LM.tex
Updates from Overleaf

% SIAM Article Template
\documentclass[r]{siamart171218}
% Packages and macros go here
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{float}
\usepackage{fullpage}
\usepackage{color}
\usepackage{enumitem}  
\usepackage{epstopdf}
\usepackage{stmaryrd}
\usepackage{tikz, subfigure}
\usepackage{booktabs}  % for tables
\usepackage{capt-of}
\usepackage{multirow}
\usepackage[displaymath, mathlines]{lineno}

@@

In this paper we will consider a 3D-1D coupled problem described by the following equations:\\
\begin{subequations}
\label{eq:pde}
\begin{align}
\label{eq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\
\label{eq:pde2}
 d_s^2 \ud + \ud - \lambda &= g &\mbox{ on } \Lambda,\\
\label{eq:coupling}
\mathcal{T}_\Lambda u - v  &=  h &\mbox{ on } \Lambda . 
\end{align}
\end{subequations}
%
Here, $u$, $\ud$,  $\lambda$ are unknowns,  $\Omega$ is a bounded domain in $\R^3$, whereas $\Lambda \subset \Omega$ is a 1D structure
parameterized in terms of $s$ and $d_s$ is the derivative with respect to $s$. 
The term $\lambda\delta_{\Lambda}$ is a Dirac measure such that 
$\int_{\Omega}\lambda(x)\delta_{\Lambda}v(x)\,\mathrm{d}x=\int_{\Lambda}\lambda(t)v(t) \,\mathrm{d}t$
for a continuous function $v$ and $\mathcal{T}_\Lambda: \Omega\rightarrow\Lambda$ is a trace operator.

From a mathematical point of view, the challenge involved in problem \eqref{eq:pde} is that neither $\mathcal{T}_\Lambda$ nor $\delta_\Lambda$ are well defined. 
That is, without extra regularity,  solutions of elliptic PDEs only have well defined traces of co-dimension one. Here, $\mathcal{T}_\Lambda$ is of co-dimension two, mapping functions defined on a domain in 3D to
functions defined along a 1D curve. Similarly, Dirac delta functions are well defined on measures of co-dimension one, but not two. Several studies \cite{DAngelo,d2012finite,d2008coupling} have remedied the well-posedness by weakening the solution concept. The approach naturally leads to non-symmetric formulations. An alternative approach is to decompose the solution into smooth and non-smooth components, where the non-smooth component may be represented in terms of Greens functions, and then consider the well-posedness of the smooth component~\cite{Peaceman1983531,secomb2004green}. Both of these approaches have successfully be analyzed in a discrete setting and implemented~\cite{}. Error estimates etc... Many more references can be mentioned. 
In this work we analyze an alternative approach, closely related to \cite{laurino_m2an}, where we exploit the fact that the $\Lambda$ is not strictly a 1D curve, but rather a very thin 3D structure with 
a cross-sectional area far below what we can resolve. With this additional assumption, we show that robustness with respect to the cross-sectional area can be restored.

The mathematical problem is motivated by the fact that many problems in geo- and biophysics are characterized by slender cylindrical structures coupled to a larger 3D body, where the characteristic transverse length scale of the slender structure is many orders of magnitude smaller than the longitudinal length.  For example, in geophysical applications the radius of wells are often of the order of 10 cm while the length may be several kilometers~\cite{gjerde2019singularity}. Similarly, in applications involving the blood flow and oxygen transport of the micro-circulation the capillary radius is a few microns, while simulations are often performed on mm to cm scale, with thousands of vessels~\cite{berg2020modelling,fang2008oxygen,gould2017capillary}. Finally, in neuro-science applications a neuron has width of a few microns, while  its length is much longer. For example, an axon of a motor neurons may be as long as a meter.  Hence,  at least 4 orders of magnitude in difference in transverse and longitudinal direction is common in both geo-physics, bio-mechanics and neuro-science. Meshes dictated by resolving the transverse length scale in 3D would then possibly lead to the order of $10^{12}$ degrees of freedom. Even if adaptive and strongly anisotropic meshes are allowed for, the computations quickly become demanding if many slender structures and their interactions are under study.  If the quantity of interest is the far field impact (relative to the transverse lengthscale) caused by processes in the slender structures, the high resolution involved in the near field of the slender structure may not be needed.

An intermediate problem, where the coupling between the 3D and 1D problems of \eqref{eq:pde} is a 2D surface, reads 
\begin{subequations}
\label{inteq:pde}
\begin{align}
\label{inteq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\
\label{inteq:pde2}
-d_s^2 \ud + \ud - E' \lambda  &= g &\mbox{ on } \Lambda,\\
\label{inteq:coupling}
T_\Gamma u - E_\Gamma v  &=  h &\mbox{ on } \Gamma
\end{align}
\end{subequations}
where $\Gamma \subset \Omega$ is a 2D surface corresponding to $\Lambda$
and $E$ is an extension operator from $\Lambda$ to $\Gamma$.

\kent{
Detail the approach and result a bit.
Mention that some stuff is 
similar to~\cite{laurino_m2an}, but for completeness and the readers convenience we repeat some basic derivations. 
}

%We address the geometrical configuration of the problem for a 3D coupled problem formulation based on from Dirichlet-Neumann interface conditions. Then, we apply a model reduction technique that transforms the problem into 3D-1D coupled PDEs. We develop and analyze a robust definition of the coupling operators form a 3D domain, $\Omega$, to 1D manifold, $\Lambda$, and vice versa. This is a non trivial objective because the standard trace operator form a domain $\Omega$ to a subset $\Lambda$ is not well posed if $\Lambda$ is a manifold of co-dimension two of $\Omega$.

%This work is based on the assumption that \emph{the transversal diameter of $\Omega_{\ominus}$ is small compared to the diameter of $\Omega$}. 
%Let $d= \max_{s \in (0,S)} \mathrm{diam}(\D(s))$ be the largest diameter of the cross sections of $\Omega_{\ominus}$, that is the transversal diameter of $\Omega_{\ominus}$.
%The central assumption of this work is that $d \ll \mathrm{diam}(\Omega)$.
%Let us now scale the domains $\Omega$ and $\Omega_{\ominus}$.
%Let $\chi_{\Omega}(\xx) = \xx / \mathrm{diam}(\Omega)$ be a scaling function and let be $\Omega_\chi = \chi_{\Omega}(\Omega), \ {\Omega_{\ominus}}_\chi =  \chi_{\Omega}(\Omega_{\ominus})$ be the scaled domains. %The previous assumption implies that for the scaled domains $\epsilon = d_\chi = d/\mathrm{diam}(\Omega)$
%is such that $0 < \epsilon \ll 1$. For simplicity of notation, and without loss of generality, form now on we will implicitly refer to the scaled domains dropping the subindex $\chi$.
%\kent{this is a good explanatory paragraph, but not sure we use the notation introduced}

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\section{Preliminaries}\label{sec:setting}

%In this preliminary section we fist provide the main notation and definitions to set up the problem.
%In particular, we address the geometry of the domain, which is equivalent to the one addressed in previous works of the authors \cite{laurino_m2an},
%and we define the functional spaces and related operators necessary for a rigorous problem formulation and analysis.
%Finally, the section ends with the formal derivation of the problem, 
%based on averaging techniques applied to a fully three-dimensional model in the geometrical setting defined before.

Let the domain $\Omega\in \R^3 $ be convex and composed of two parts, 
$\Omega_{\ominus}$ and $\Omega_{\oplus}:=\Omega\setminus\overline{\Omega}_{\ominus}$. 
Let 
$\Omega_{\ominus}$ be a \emph{generalized cylinder}, c.f.  \cite{MR1940257}, 
that is; the swept volume of a two dimensional set, $\partial D$, moved along a curve, $\Lambda$, in the three-dimensional domain, $\Omega$, see for Figure \ref{fig1} for an illustration. 
In detail, the curve 
$\Lambda = \{\boldsymbol \lambda(s), \ s\in(0,S)\}$,  where 
$\boldsymbol \lambda(s) = [\xi(s), \tau(s), \zeta(s)], \ s\in(0,S)$ is a $\mathcal{C}^2$-regular curve in the three-dimensional domain $\Omega$.
For simplicity, let us assume that $\|\boldsymbol \lambda^\prime(s)\|=1$ such that the arc-length and the coordinate $s$ coincide.
Further, let $\mathcal{D}(s) = [x(r,t),y(r,t)]: (0,R(s))\times(0,T(s)) \rightarrow \R^2$ be a parametrization of the cross section
and $\Gamma$ be the lateral surface of $\Omega_{\ominus}$, i.e.
$\Gamma=\{ \ \partial D(s) \  | \ s \in \Gamma\}$,
while the upper and lower faces of $\Omega_{\ominus}$ belong to $\partial\Omega$. 
We will assume that $\Omega_{\ominus}$ crosses $\Omega$ from si
de to side.
According to the geometrical setting, we will denote with $v,\,v_\oplus,\,v_\ominus,\,\vd$,
functions defined on $\Omega,\,\Omega_{\oplus},\,\Omega_{\ominus},\,\Lambda$, respectively.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.3\textwidth]{domain.pdf}
\end{center}
\caption{Geometrical setting of the problem}
\label{fig1}
\end{figure}

%Let $\TT,\NN,\BB$ be the Frenet frame related to the curve.
%Let $\mathcal{D}(s) = [x(r,t),y(r,t)]: (0,R(s))\times(0,T(s)) \rightarrow \R^2$ be a parametrization of the cross section.
%Let us assume that $\mathcal{D}(s)$ is convex for any $s \in (0,S)$. The cross section can change size along $\Lambda$ but not shape.
%Let us also parametrize the boundary of the cross section as $\partial\mathcal{D}(s) = [\partial %x(r,t),\partial y(r,t)]: (0,R(s))\times(0,T(s)) %\rightarrow \R^2$,
%and let us assume that $\partial\mathcal{D}(s)$ is a %piecewise $\mathcal{C}^2$-regular curve.
%Then, the generalized cylinder $\Omega_{\ominus}$ %can be defined as follows
%\begin{equation*}
%\Omega_{\ominus} = \{ \boldsymbol \lambda(s) + %x(r,t) \NN(s) + y(r,t) \BB(s), \ r \in (0,R(s)),\ s %\in (0,S), \ t \in (0,T(s))\}\,,
%\end{equation*}
%and the lateral boundary is
%\begin{equation*}
%\Gamma = \{ \boldsymbol \lambda(s) + \partial x(r,t) %\NN(s) + \partial y(r,t) \BB(s), \ r \in (0,R(s)),\ %s \in (0,S), \ t \in (0,T(s))\}\,.
%\end{equation*}
%Let $|\cdot|$ denote the Lebesgue measure of a set. %If $|\mathcal{D}(0)|, \ |\mathcal{D}(S)| >0$ then %$\Omega_{\ominus}$ has \emph{top} and \emph{bottom} %boundaries,
%which belong to $\partial\Omega$.

% >>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
%\subsection{Functional setting and preliminary results}

Let $D$ be a generic regular bounded domain in $\R^3$
and $X$ be a Hilbert space defined on $D$ then  $(\cdot,\cdot)_X$ and $\|\cdot\|_X$ denote the inner product and norm of $X$, respectively.
The duality pairing between the $X$ and its dual $X^*$ is denoted as $\langle\cdot,\cdot\rangle$.
Let $(\cdot,\cdot)_{L^2(D)}$,  $(\cdot,\cdot)_D$ or simply $(\cdot,\cdot)$ be the $L^2(D)$ inner product on $D$.
We will use the standard notation $H^q(D)$ to denote the Sobolev space of functions on $D$ with all derivatives up to the order $q$ in $L^2(D)$.
The corresponding norm is $\|\cdot\|_{H^q(D)}$ and the seminorm is $\vert\cdot\vert_{H^q(D)}$.
The space $H^q_0(D)$ represents the closure in $H^q(D)$ of smooth functions with compact support in $D$.

@@

We will frequently used inner products and norms that are weighted, such if $X$ is a Hilbert space and $w$ is a positive number, then the
weighted inner product is denoted 
\[
(u,v)_{X,w} = (w \, u,  v)_X
\]
with a corresponding norm $\|\cdot\|_{X,w}$. Finally, 
$|\cdot|$ denote the Lebesgue measure of a set, i.e. $|D(0)|$ is the cross-sectional area of the bottom of the cylinder.
We will in the following assume that the cross-sectional area $D(s)$ is constant, but remark that this assumption is only made to simplify the notation. 
\kent{I wrote this comment for simplicity. Otherwise, we need to be a bit more careful with the definition of the inner product above}

Central in our analysis are the transverse averages $\avrc{w},\,\avrd{w}$ defined as, 
\begin{gather*}
\avrc{w}(s) = |\partial\mathcal{D}(s)|^{-1} \int_{\partial\mathcal{D}(s)} w d\gamma \quad \mbox{and} \quad 
\avrd{w}(s) = |\mathcal{D}(s)|^{-1} \int_{\mathcal{D}(s)} w d\sigma, 
\end{gather*}
where $d\omega, \ d\sigma, \ d\gamma$ are the generic volume, surface and curvilinear Lebesgue measures.
Clearly, 
\begin{gather*}
\int_{\Omega_{\ominus}} w d\omega 
= \int_\Lambda \int_{\mathcal{D}(s)} w d\sigma ds
= \int_\Lambda |\mathcal{D}(s)|\avrd{w}(s) ds\, \ \mbox{and} \ 
\int_{\partial\Omega_{\ominus}} w d\sigma 
= \int_\Lambda \int_{\partial\mathcal{D}(s)} w d\gamma ds
= \int_\Lambda  |\partial \mathcal{D}(s)| \avrc{w}(s) ds\,. 
\end{gather*}
Analogously, for functions defined on $\Lambda$ and $\Omega_\ominus$ respectively, 
we define as $d_s$ and $\partial_s$ the ordinary and partial derivative with respect to the arclength.

The operator obtained from a combination of the average operator $\avrc{(\cdot)}$ with the trace on $\Gamma$ will be denoted with $\mtrace = \avrc{(\cdot)} \circ \trace$,
as it maps functions on $\Omega$ to functions on $\Lambda$.
Further, let the extension operator $\mathcal{E}_{\Gamma}: H^{\frac 1 2}_{00}(\Lambda) \rightarrow H^{\frac 1 2}_{00}(\Gamma)$ be defined such that
 $(\mathcal{E}_{\Omega_{\ominus}}\vd)(x)=\vd(s), \quad \forall x\in \partial D(s)$.
Then, 
\begin{equation}\label{eq:ext}
\langle \mtrace u, \vd \rangle_{\Lambda, |\DD|} 
= \int _{\Lambda} |\DD| \left( \frac{1}{|\DD|} \int_{\DD} \trace u \, d\gamma \right) \vd \, ds 
= \langle \trace u, \ext \vd\rangle_{\Gamma}\,,
\end{equation} 
which shows that the average and the uniform extension operator commute with the trace.

@@

With the above notation we are now able to formulate the precise weak formulations of the problems 
\eqref{eq:pde} and \eqref{inteq:pde}. 
The weak formulation of \eqref{eq:pde}, which we will call the \textbf{Problem 3D-1D-1D} reads: 
Find $u \in H^1_0(\Omega),\ \ud \in H^1_0(\Lambda), \ \ld \in H^{-\frac 1 2}(\Lambda)$, such that
\begin{subequations}\label{eq:problem2}
\begin{align}
&(u,v)_{H^1(\Omega)} + (\ud,\vd)_{H^1(\Lambda),|\D|} 
+  \langle \mtrace v - \vd, \ld \rangle_{\Lambda, |\DD|} &&
\\
\nonumber
&\qquad\qquad= (f,v)_{L^2(\Omega)} +  (\avrd{g},V)_{L^2(\Lambda),|\D|}
\quad && \forall v \in H^1_0(\Omega), \ \vd \in H^1_0(\Lambda)\,,
\\
& \langle \mtrace u -   \ud, \md \rangle_{\Lambda,|\DD|} = 0
\quad &&\forall \md \in H^{-\frac 1 2}(\Lambda)\,,
\end{align}
\end{subequations}
Furthermore, the weak formulation of \eqref{inteq:pde}, refered to as \textbf{Problem 3D-1D-2D}, reads: 
Find $u \in H^1_0(\Omega), \ \lambda \in H^{-\frac 1 2}(\Gamma ),\ \ud \in H^1_0(\Lambda)$, such that
\begin{subequations}\label{eq:problem1}
\begin{align}
&(u,v)_{H^1(\Omega)} + (\ud,\vd)_{H^1(\Lambda),|\D|} 
+ \langle \trace v  - \ext \vd, \lambda \rangle_\Gamma &&
\\
\nonumber
&\qquad= (f,v)_{L^2(\Omega)} +  (\avrd{g},\vd)_{L^2(\Lambda),|\D|}
\quad &&\forall v \in H^1_0(\Omega), \ \vd \in H^1(\Lambda)\,,
\\
& \langle \trace u - \ext \ud , \mu \rangle_\Gamma = 0
\quad && \forall \mu \in H^{-\frac 1 2}(\Gamma)\,.
\end{align}
\end{subequations}

\section{Derivation of the model}
Below we detail the steps performed in order to formulate \ref{eq:problem1} and \ref{eq:problem2} precisely. 
The derivations are similar~\cite{laurino_m2an},  but here we derive the system including the Lagrange multiplier.

\subsection{Topological model reduction}

We consider the problem arising from \emph{Dirichlet-Neumann} conditions. Find $\up,\uf$ s.t.:
\begin{subequations}\label{eq:dirneu}
\begin{align}
\label{eq:dirneu1}
- \Delta \up  + \up &= f  && \text{ in } \Omega_{\oplus},\\
\label{eq:dirneu2}
- \Delta \uf  + \uf &= g  && \text{ in } \Omega_\ominus,\\
\label{eq:dirneu3}
-\nabla \uf \cdot \nn_{\ominus} &= -\nabla \up \cdot \nn_{\ominus}  && \text{ on } \Gamma,\\
\label{eq:dirneu4}
\uf &= \up && \text{ on }  \Gamma,\\
\label{eq:dirneu5}
\up &= 0 && \text{ on } \partial \Omega\,.
\end{align}
\end{subequations}

We observe that the coupling constraints defined on $\Gamma$ involve essential or strong conditions.
The strong conditions will, however, be enforced weakly by using the method of Lagrange multipliers~\cite{MR359352}.
%We remark that this case is substantially different from the one previously addressed in \cite{laurino_m2an}.

>>>>

% SIAM Article Template
\documentclass[r]{siamart171218}
% Packages and macros go here
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{array}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{float}
%\usepackage{fullpage}
\usepackage{color}
\usepackage{enumitem}  
\usepackage{epstopdf}
\usepackage{stmaryrd}
\usepackage{standalone}
\usepackage{tikz, subfigure}

\usetikzlibrary{shapes, snakes, patterns, arrows}
\usetikzlibrary{calc}
\usepackage{pgfplots}
\usepgfplotslibrary{fillbetween}

\usepackage{booktabs}  % for tables
\usepackage{capt-of}
\usepackage{multirow}
\usepackage[displaymath, mathlines]{lineno}

@@

In this study we consider coupled partial differential equations on domains with mixed dimensionality, in particular we address the  3D-1D case. The mathematical structure of such problems can be represented by the following formal equations:
\begin{subequations}
\label{eq:pde}
\begin{align}
\label{eq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\
\label{eq:pde2}
 d_s^2 \ud + \ud - \lambda &= g &\mbox{ on } \Lambda,\\
\label{eq:coupling}
\mathcal{T}_\Lambda u - v  &=  h &\mbox{ on } \Lambda . 
\end{align}
\end{subequations}
Problem \eqref{eq:pde} can be described as an example of \emph{mixed dimensional PDEs}.
Here, $u$, $\ud$,  $\lambda$ are unknowns,  $\Omega$ is a bounded domain in $\R^3$, whereas $\Lambda \subset \Omega$ is a 1D structure
parameterized in terms of $s$ and $d_s$ is the derivative with respect to $s$. 
The term $\lambda\delta_{\Lambda}$ is a Dirac measure such that 
$\int_{\Omega}\lambda(x)\delta_{\Lambda}v(x)\,\mathrm{d}x=\int_{\Lambda}\lambda(t)v(t) \,\mathrm{d}t$
for a continuous function $v$ and $\mathcal{T}_\Lambda: \Omega\rightarrow\Lambda$ is a suitable restriction operator from 3D to 1D.

Using models based on mixed dimensional PDEs is motivated by the fact that many problems in geo- and biophysics are characterized by slender cylindrical structures coupled to a larger 3D body, where the characteristic transverse length scale of the slender structure is many orders of magnitude smaller than the longitudinal length.  For example, in geophysical applications the radius of wells are often of the order of 10 cm while the length may be several kilometers~\cite{Peaceman1978183, Peaceman1983531}. Similarly, in applications involving the blood flow and oxygen transport of the micro-circulation the capillary radius is a few microns, while simulations are often performed on mm to cm scale, with thousands of vessels~\cite{berg2020modelling,fang2008oxygen,gould2017capillary, secomb2004green}. Finally, in neuro-science applications a neuron has width of a few microns, while  its length is much longer. For example, an axon of a motor neurons may be as long as a meter.  Hence,  at least 4 orders of magnitude in difference in transverse and longitudinal direction is common in both geo-physics, bio-mechanics and neuro-science. Meshes dictated by resolving the transverse length scale in 3D would then possibly lead to the order of $10^{12}$ degrees of freedom. Even if adaptive and strongly anisotropic meshes are allowed for, the computations quickly become demanding if many slender structures and their interactions are under study.

From a mathematical standpoint, the challenge involved in problem \eqref{eq:pde} is that neither $\mathcal{T}_\Lambda$ nor $\delta_\Lambda$ are well defined. That is, without extra regularity,  solutions of elliptic PDEs only have well defined traces of co-dimension one. Here, $\mathcal{T}_\Lambda$ is of co-dimension two, mapping functions defined on a domain in 3D to functions defined along a 1D curve. 
The challenge of coupling PDEs on domains with high dimensionality gap has recently attracted the attention of many researchers. The sequence of works by D'Angelo, \cite{DAngelo,d2012finite,d2008coupling} have remedied the well-posedness by weakening the solution concept. The approach naturally leads to non-symmetric formulations. An alternative approach is to decompose the solution into smooth and non-smooth components, where the non-smooth component may be represented in terms of Green functions, and then consider the well-posedness of the smooth component \cite{gjerde2019singularity}. The numerical approximation of such equations has been also studied in a series of works. The consistent derivation of numerical approximations schemes for PDEs in mixed dimension is addressed in \cite{Nordbotten}. Concerning approximability,  elliptic equations with Dirac sources represent an effective prototype case that has been addressed in \cite{Bertoluzza201897,koppl2016local,koppl2014optimal}, where the optimal a-priori error estimates for the finite element approximation are derived. Furthermore, the interplay between the mathematical structure of the problem and solvers, as well as preconditioners for its discretization has been studied in details  in \cite{Kuchta2016B962} for the solution of 1D differential equations embedded in 2D, and more recently extended to the 3D-1D case in \cite{KMM2}.

Stemming from this literature, in this work we adopt and analyze a different approach, closely related to \cite{KVWZ,laurino_m2an}. That is, we exploit the fact that the $\Lambda$ is not strictly a 1D curve, but rather a very thin 3D structure with a cross-sectional area far below what can be resolved. With this additional assumption, we show that robustness with respect to the cross-sectional area can be restored. The major novelty with respect to the previous works is that we address essential type coupling conditions, namely Dirichlet-Neumann conditions, rather than natural type ones, such as the Neumann-Robin or Robin-Robin cases. These coupling conditions pose additional difficulties as the conditions 
are not a natural part of the weak formulation of the problem. We overcome this difficulty by  resorting to a weak formulation of the Dirichlet-Neumann coupling conditions across dimensions by using Lagrange multipliers.

%>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
\section{Preliminaries}\label{sec:setting}

Let the domain $\Omega\in \R^3 $ be convex and composed of two parts, 
$\Omega_{\ominus}$ and $\Omega_{\oplus}:=\Omega\setminus\overline{\Omega}_{\ominus}$. 
Let 
$\Omega_{\ominus}$ be a \emph{generalized cylinder}, c.f.  \cite{MR1940257}, 
that is; the swept volume of a two dimensional set, $\partial D$, moved along a curve, $\Lambda$, in the three-dimensional domain, $\Omega$, see for Figure \ref{fig1} for an illustration. 
In detail, the curve 
$\Lambda = \{\boldsymbol \lambda(s), \ s\in(0,S)\}$,  where 
$\boldsymbol \lambda(s) = [\xi(s), \tau(s), \zeta(s)], \ s\in(0,S)$ is a $\mathcal{C}^2$-regular curve in the three-dimensional domain $\Omega$.
For simplicity, let us assume that $\|\boldsymbol \lambda^\prime(s)\|=1$ such that the arc-length and the coordinate $s$ coincide.
Further, let $\mathcal{D}(s) = [x(r,t),y(r,t)]: (0,R(s))\times(0,T(s)) \rightarrow \R^2$ be a parametrization of the cross section
and $\Gamma$ be the lateral surface of $\Omega_{\ominus}$, i.e.
$\Gamma=\{ \ \partial D(s) \  | \ s \in \Gamma\}$,
while the upper and lower faces of $\Omega_{\ominus}$ belong to $\partial\Omega$. 
We assume that $\Omega_{\ominus}$ crosses $\Omega$ from side to side.
According to the geometrical setting, we will denote with $v,\,v_\oplus,\,v_\ominus,\,\vd$,
functions defined on $\Omega,\,\Omega_{\oplus},\,\Omega_{\ominus},\,\Lambda$, respectively.

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.25\textwidth,angle=-90]{./graphics/domain.tex}
\end{center}
\caption{Geometrical setting of the problem}
\label{fig1}
\end{figure}
\kent{I think the symbols in the Figure should be bigger}

Let $D$ be a generic regular bounded domain in $\R^3$ and $X$ be a Hilbert space defined on $D$ then  $(\cdot,\cdot)_X$ and $\|\cdot\|_X$ denote the inner product and norm of $X$, respectively.
The duality pairing between the $X$ and its dual $X^*$ is denoted as $\langle\cdot,\cdot\rangle$.
Let $(\cdot,\cdot)_{L^2(D)}$,  $(\cdot,\cdot)_D$ or simply $(\cdot,\cdot)$ be the $L^2(D)$ inner product on $D$.
We use the standard notation $H^q(D)$ to denote the Sobolev space of functions on $D$ with all derivatives up to the order $q$ in $L^2(D)$.
The corresponding norm is $\|\cdot\|_{H^q(D)}$ and the seminorm is $\vert\cdot\vert_{H^q(D)}$.
The space $H^q_0(D)$ represents the closure in $H^q(D)$ of smooth functions with compact support in $D$.

@@

We will frequently use inner products and norms that are weighted, such that if $X$ is a Hilbert space and $w$ is a positive integrable function, then the
weighted inner product is denoted 
\[
(u,v)_{X,w} = (w \, u,  v)_X
\]
\kent{
I think we need to be careful when we write this. A reader may believe
that 
\[
(u,v)_{1,w} = (\nabla (w \, u),  \nabla v)_1
\]
I will fix this, but I'd like a chat first.
}
with a corresponding norm $\|\cdot\|_{X,w}$. Finally, $|\cdot|$ denote the Lebesgue measure of a set, i.e. $|D(0)|$ is the cross-sectional area of the bottom of the cylinder. In the following we assume that the cross-sectional area $\mathcal{D}(s)$ is constant, namely $\mathcal{D}(s)=\mathcal{D}$, but remark that this assumption is only made to simplify the notation. Central in our analysis are the transverse averages $\avrc{w},\,\avrd{w}$ defined as, 
\begin{gather*}
\avrc{w}(s) = |\partial\mathcal{D}(s)|^{-1} \int_{\partial\mathcal{D}(s)} w d\gamma \quad \mbox{and} \quad 
\avrd{w}(s) = |\mathcal{D}(s)|^{-1} \int_{\mathcal{D}(s)} w d\sigma, 
\end{gather*}
where $d\omega, \ d\sigma, \ d\gamma$ are the generic volume, surface and curvilinear Lebesgue measures.
Clearly, 
\begin{gather*}
\int_{\Omega_{\ominus}} w d\omega 
= \int_\Lambda \int_{\mathcal{D}(s)} w d\sigma ds
= \int_\Lambda |\mathcal{D}(s)|\avrd{w}(s) ds\, 
\\
\int_{\partial\Omega_{\ominus}} w d\sigma 
= \int_\Lambda \int_{\partial\mathcal{D}(s)} w d\gamma ds
= \int_\Lambda  |\partial \mathcal{D}(s)| \avrc{w}(s) ds\,. 
\end{gather*}
Analogously, for functions defined on $\Lambda$ and $\Omega_\ominus$ respectively, 
we let  $d_s$ and $\partial_s$ be the ordinary and partial derivative with respect to the arclength.

The operator obtained from a combination of the average operator $\avrc{(\cdot)}$ with the trace on $\Gamma$ will be denoted with $\mtrace = \avrc{(\cdot)} \circ \trace$, as it maps functions on $\Omega$ to functions on $\Lambda$.
Further, let the extension operator $\mathcal{E}_{\Gamma}: H^{\frac 1 2}_{00}(\Lambda) \rightarrow H^{\frac 1 2}_{00}(\Gamma)$ be defined such that $(\mathcal{E}_{\Omega_{\ominus}}\vd)(x)=\vd(s)$, for any $x\in \DD(s)$. Then, the following identity shows that the transversal uniform extension operator is the inverse of the transversal average.

\begin{equation}\label{eq:ext}
\langle \mtrace u, \vd \rangle_{\Lambda, |\DD|} 
= \int _{\Lambda} |\DD| \left( \frac{1}{|\DD|} \int_{\DD} \trace u \, d\gamma \right) \vd \, ds 
= \langle \trace u, \ext \vd\rangle_{\Gamma}\,,
\end{equation}

@@

With the above notation we are now able to formulate the precise weak formulations of the problems \eqref{eq:pde} and \eqref{inteq:pde}. The weak formulation of \eqref{eq:pde}, which we will call the \textbf{Problem 3D-1D-1D} reads:  Find $u \in H^1_0(\Omega),\ \ud \in H^1_0(\Lambda), \ \ld \in H^{-\frac 1 2}(\Lambda)$, such that
\begin{subequations}\label{eq:problem2}
\begin{align}
(u,v)_{H^1(\Omega)} 
+  \langle \mtrace v, \ld \rangle_{\Lambda, |\DD|} &= 
(f,v)_{L^2(\Omega)} &\forall v \in H^1_0(\Omega)\, ,\\
(\ud,\vd)_{H^1(\Lambda),|\D|} - \langle  \vd, \ld \rangle_{\Lambda, |\DD|} 
&=  (\avrd{g},V)_{L^2(\Lambda),|\D|}
&\forall \vd \in H^1_0(\Lambda)\, ,\\
\langle \mtrace u -   \ud, \md \rangle_{\Lambda,|\DD|} &= 0
&\forall \md \in H^{-\frac 1 2}(\Lambda)\,,
\end{align}
\end{subequations}

In addition to the 3D-1D-1D problem we will also consider an 
intermediate problem where the 3D and 1D problems are coupled at an
intermediate 2D surface  encapsulating the 1D structure. The strong form is:  
\begin{subequations}
\label{inteq:pde}
\begin{align}
\label{inteq:pde1}
  -\Delta u + u + \lambda \delta_{\Lambda} &= f &\mbox{ in } \Omega,\\ 
\label{inteq:pde2}
 d_s^2 \ud + \ud - \lambda &= g &\mbox{ on } \Lambda,\\
\label{inteq:coupling}
\mathcal{T}_\Lambda u - v  &=  h &\mbox{ on } \Gamma . 
\end{align}
\end{subequations}

The corresponding weak formulation of \eqref{inteq:pde}, refered to as the \textbf{Problem 3D-1D-2D}, reads: 
Find $u \in H^1_0(\Omega), \ \ud \in H^1_0(\Lambda), \ \lambda \in H^{-\frac 1 2}(\Gamma ) $ such that
\begin{subequations}\label{eq:problem1}
\begin{align}
(u,v)_{H^1(\Omega)} + \langle \trace v, \lambda \rangle_\Gamma &= (f,v)_{L^2(\Omega)} &\forall v \in H^1_0(\Omega)\, ,
\\
 (\ud,\vd)_{H^1(\Lambda),|\D|}  -  \langle \ext \vd, \lambda \rangle_\Gamma
&=  (\avrd{g},\vd)_{L^2(\Lambda),|\D|} &\forall\vd \in H^1(\Lambda)\, ,\\
\langle \trace u - \ext \ud , \mu \rangle_\Gamma &= 0 &\forall \mu \in H^{-\frac 1 2}(\Gamma)\,.
\end{align}
\end{subequations}

\begin{lemma}\label{lemma:H12norm}
Let $\Gamma$ be a tensor product domain, $\Gamma= (0,X) \times (0,Y)$. For any regular $u(x,y)$ in $\Gamma$, let $\avrc{u}(x)=\frac 1Y \int _0^Y u(x,y)\, dy$. Then, for any $u\in H_{00}^{\frac 12}(\Gamma)$, then $\avrc{u}(x)\in H_{00}^{\frac 12}((0,X))$. 
Moreover, if $u(x,y)\in H^{\frac 12}_{00}(\Gamma)$ is constant with respect to $y$, namely $u(x,y)=u(x)$, then 
\begin{equation*}
\|u\|_{H^{\frac 12}_{00}(\Gamma)}=Y \|u\|_{H^{\frac 12}_{00}(0,X)}.
\end{equation*}
\end{lemma}
The proof of \label{lemma:H12norm} is based on the representation of fractional norms in terms of the spectrum of the Laplace operator and subsequent standard arguments in harmonic analysis. The full proof is reported in the appendix for the sake of clarity.

%\begin{corollary}
%Let $\Gamma$ be the lateral surface of a prism with $N$ faces $\Gamma_i$, $i=1, \dots, N$, with $\Gamma_i=(0,X_i)\times(0,S)$. For any $u\in H^{\frac 12}_{00}(\Gamma)$, let $\avrc{u}(s)= \frac{1}{\sum_i{|X_i|}}\sum_{i=1}^N \int_0^{X_i} u(x,s)\, dx$. Then $\avrc{u}\in H^{\frac 12}_{00}(0,S )$ {\color{red} and there exists a constant $C_\Gamma$ such that
%\begin{equation*}
%\|\avrc{u}\|_{H^{\frac 12}(0,S)} \leq C_\Gamma \|u\|_{H^{\frac 12}(\Gamma)}.
%\end{equation*}
%}
%\end{corollary}
